Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-10-08 exp1/present2013.tex
fractional experiments
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@655 43ef1190-90fd-4dd7-a769-d0121f02b80e

>>>>

\subsection{Fractional Factorial Experiments}

\begin{frame}{Experimental Economy}

  \begin{itemize}
  \item Full factorial experimental goes as $2^n$.
  \item SPARQL Endpoint evaluation is inherently complex, with many
    possible factors.
  \item May reduce the size of the experiment with Fractional
    Factorial Experiments.
  \item The cost is explanatory power due to \emph{aliasing}.
  \item E.g. we cannot tell the difference of an effect of
    ``TripleC::BGPComp'' from ``Machine::Range''
  \item Easy in R to declare which effects must not be aliased.
  \item We have made two designs: One 32-run and one 64-run.
  \end{itemize}

\end{frame}
\note{ That is to say, a detected increase in
run time from larger Basic Graph Patterns for large databases, can
also be explained by a less powerful machine that is worse at
evaluating FILTER clauses with ranges. }

\begin{frame}{32-run experiment}
\begin{figure}[ht!]
  \centerline{%
    \includegraphics[height=0.85\textheight,width=.9\textwidth]{frac32normal.pdf}}
  \label{fig:frac32normal}
\end{figure}

\end{frame}
\note{The total variance is great, due to few runs, so few conclusions
  can be drawn}
\note{Implementation is good, and Union is hard}
\note{In the paper, we also do a hypothesis test that returns  $p =
  0.00098$. Still very low}
\note{Very rough conclusions; useful in a CI system}

\begin{frame}{64-run experiment}
\begin{figure}[ht!]
  \centerline{%
    \includegraphics[height=0.85\textheight,width=0.96\textwidth]{frac64normal.pdf}}
\end{figure}

\end{frame}
\note{Have two Lenth criterions, $\alpha=0.15$ and crosses at
  $\alpha=0.05$. ($\alpha$ is the probability of rejecting $H_0$ when
  it is true)}

\begin{frame}{Analysis of 64-run experiment}

  \begin{itemize}
  \item Find most of the same significant effects as the full experiment.
  \item However, we see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$.
  \item ``Machine'', ``TripleC'' and ``Range'' are inactive, we can
    perform hypothesis test as above, and find $p = 5.9 \cdot 10^{-6}$.
  \end{itemize}
\end{frame}
\note{We see that ``Implement:TripleC'' emerges only for
  $\alpha=0.15$, which is important to better in real studies.}
\note{We have shown that how the experiment could be set up, we have
  shown the analysis and a hypothesis test.}
\note{Should this simplistic experiment stand up to scrutiny?}

\subsection{Tearing it all down}

\begin{frame}{Fractional experiment with more data}
\begin{figure}[ht!]
  \centerline{%
  \includegraphics[height=0.85\textheight,width=.9\textwidth]{frac64hugenormal.pdf}}
\end{figure}

\end{frame}
\note{``Machine'' is a broad factor, good idea to group to manage
  complexity, can be broken down later}
\note{``TripleC:Machine'' and ``TripleC'' highly significant, implies
  the factors were wrongly set.}
\note{This invalidates the three preceeding experiments. We got strong
indications early on.}

\section{Conclusion}
